---
title: "Getting Started with DPrivStats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with DPrivStats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(DPrivStats)
set.seed(42)
```

This vignette walks through a basic differentially private (DP) analysis
workflow: loading sensitive microdata, managing a privacy budget, releasing
DP descriptive statistics, and running a DP hypothesis test.

## Data

We use a simulated census-like microdata set bundled with the package:

```{r}
data(example_microdata)
head(example_microdata)
nrow(example_microdata)
```

## Privacy budget

All releases draw on a shared privacy budget:

```{r}
budget <- new_privacy_budget(epsilon = 3.0, delta = 1e-6, composition = "rdp")
budget
```

## DP descriptive statistics

```{r}
income_bounds <- c(0, 500000)
m <- dp_mean(example_microdata$income, epsilon = 0.5,
             bounds = income_bounds, mechanism = "laplace")
m
```

The estimate carries its privacy cost and sensitivity with it:

```{r}
m$sensitivity
```

A DP median via the exponential mechanism, and a DP histogram:

```{r}
dp_median(example_microdata$income, epsilon = 0.5, bounds = income_bounds,
          n_bins = 200)$estimate

hist_fit <- dp_histogram(example_microdata$income, epsilon = 1.0,
                         breaks = seq(0, 500000, by = 100000),
                         normalize = TRUE)
round(hist_fit$estimate, 4)
```

## DP t-test

Do the two regions differ in income?

```{r}
sub <- subset(example_microdata, region %in% c("North", "South"))
tt <- dp_t_test(x = sub$income[sub$region == "North"],
                y = sub$income[sub$region == "South"],
                epsilon = 1.0, bounds = income_bounds)
tt
```

## Budget accounting

```{r}
budget <- spend(budget, 0.5, description = "DP mean income")
budget <- spend(budget, 0.5, description = "DP median income")
budget <- spend(budget, 1.0, description = "DP histogram")
budget <- spend(budget, 1.0, description = "DP t-test")
budget
```
